---
title: "Exploratory Numeric Analysis"
output: github_document
---
```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

## all plots i make will have the viridis color palette
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


## Load the weather data

rnoaa package allows you to access data from online
floor_date -> round down
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## `group_by`

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  ungroup()
```

## counting things
How many _________ exist?

Count month / name observations
n_obs is the variable created to define the function n()
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

We can use `count()`
Does the same thing as above
```{r}
weather_df %>% 
  count(month, name = "n_obs")
```

 **NEVER** USE BASE R'S `TABLE` function
The result of this is not a dataframe
YOu have to pull a variable out of a datafram to use the table function
```{r}
weather_df %>% 
  pull(month) %>% 
  table()
```
 
Other helpful counters
n_obs - number of observations in that month
n_days - number of unique days in that month (how many dates in that month are distinct)
```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(n_obs = n(),
            n_days = n_distinct(date))
```

## A digression on 2X2 tables 
case_when is kind of like if/then
TRUE ~ "" is just in case some observations don't fit into the categories defined

Data that goes into a 2X2 table
```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
    )
  ) %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

To actually make it look like a 2X2 table
```{r}

weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
    )
  ) %>% 
  janitor::tabyl(name, cold)
```

## General summaries 

You can do lots of summaries

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE), 
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )
```

This is a dataframe!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE), 
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

Suppose you want to summarize many columns in the same way

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))

```

Reminder: sometimes your results are easier to read in another format
We can use "pivot wider" because we are doing these calculations in a dataframe
pivot wider is easier to read
"knitr::kable()" helps format info so it is easier to read (creates a table)
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
knitr::kable(digits = 1)
```

## `group_by` and `mutate`

mutate added a column to the existing dataframe

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    center_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y = center_tmax, color = name)) +
  geom_point()
```


What about window functions?

Rank function - ranks the values
Min_rank ranks coldest as 1 and increases
if you add descending 
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(temp_rank = min_rank(tmax))
```

lag - shift variables by 1
good for calculating change

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(temp_change = tmax - lag(tmax)) %>% 
  summarize(
    temp_change_max = max(temp_change, na.rm = TRUE),
    temp_change_sd = sd(temp_change, na.rm = TRUE)
  )
```

##Quick Note

summarize only gets you so far
