---
title: "Visualization Part 1"
output: github_document
---
```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the weather data
rnoaa package allows you to access data from online
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatter plots!!!

Create first scatter plot
Have to add "geom_point" to add points into the plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

New approach, same plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save and edit an object
Work flow that is least used. 
```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()

```

## Advanced scatter plot

Start with the same one and make it fancy

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth( se = FALSE)
```

What about the aes placement?
Move color down the geom point --> plot looks the without geom smoot
with geom smooth, there is one solid line (color only applies to the scatter plot)
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth()
```

Lets facet things!!
Facet grid lets you define number of rows and columns 
  . means nothing defines rows
  ~ name means name variable defines columns
alpha = reduces the transparency of the points
  can map alpha to a variable to create a gradient
size = size of the line
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, color = name)) +
  geom_point() +
  geom_smooth( se = FALSE, size = 1) +
  facet_grid(. ~name)
```

Lets combine some elements and try a new plot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.1) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~name)
```

## small notes

How many geoms have to exist?
  You can have as many as you want 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

You can use a neat geom!
  hex - hexagons
  geom_density2d( -> density plot)
  For large datasets!!
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d() +
  geom_point(alpha = 0.03)
```

## Univariate plots

One variable at a time

Histograms are really great

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Color and histograms

color = name works but it outlines the bars, doesn't change the color of the bar. Also the bars overlap each other

geom_histogram(position = "dodge") moves bars to the left and the right and doesn't stack them

fill = name fills the bar instead of outlining them

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~name)
```

Let's try a new geometry
  geom_density is a histogram that has been smoothed out around the edges

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.4) 
```

Boxplots!!

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

Trendy plots :-)

Violin plot - looks like a violin
  cross between a boxpot and and a density plot
  can add stat_summary to get different stats
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) + 
  geom_violin(alpha = 0.5) +
  stat_summary(fun = "median")
```

Ridge plot - most popular of 2017
  density curves are stacked instead of on top of each other
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```


