---
title: "Visualization Part 2"
output: github_document
---
```{r setup}
library(tidyverse)
library(patchwork)
```

## Load the weather data
rnoaa package allows you to access data from online
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Remember this plot?


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) 
```

## Labels

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  )
```

## Scales

Start with the same plot
"trans = sqrt/log" changes the type of plot
"position = " can move the axis around the graph

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0 ,15),
    labels = c("-15 C", "0 C", "15 C")
  ) +
  scale_y_continuous(
    position = "right"
  )
```

Look at color scales 
"scale_color_hue(h = c(100,300))"
  hues go from 0 to 360 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
 scale_color_hue(
   name = "Location",
   h = c(100,300))
```

Other color option (virdis package)
This package assumes that the value you're putting in is on a continuous scale --> we need to tell it that the variable is discrete

This is a good color palette to use
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
 viridis::scale_color_viridis(
   name = "Location",
   discrete = TRUE)

```

## Themes

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
 viridis::scale_color_viridis(
   name = "Location",
   discrete = TRUE) +
  theme(legend.position = "bottom")
```

Change the overall grey theme
theme_bw - changes background the white and outline to black
theme_minimal - takes away outline
theme_classic - no gridline
can use "ggthemes::" to explore the other themes 

To apply 2 themes, you cannot stack the theme function twice
  update overall theme first and then tweek things like legend position
```{r}

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
 viridis::scale_color_viridis(
   name = "Location",
   discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Setting options

Generally this is code that we run at the very beginning of the rmd file
fig.asp is aspect ratio of width to height

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chuck$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

## all plots i make will have the viridis color palette
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

