---
title: "Data Import"
output: github_document
---
```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## Read in some data

Read in the litters dataset. 
"read_csv" exisits in the readr package in tidyverse
Since we're working within the "data_wrangling_i" project, we can create a relative path starting with that folder
  "./" means start with the project folder
  
Also will update names in df because they aren't in good form (clean_names)
  Doing "janitor::" means clean_names is in that package. But since    it's the only function in that package we need, we don't need to     load the entire package

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## Take a look at the data

Printing in the console
```{r}
litters_df
```

Prints first 6 observations
```{r}
head(litters_df)
```

Print last 6 observations 
```{r}
tail(litters_df)
```

Some descriptive info
```{r}
skimr::skim(litters_df)
```

Use view(litter_df) in console to open up a new tab of the data set. Run this in the console because it makes knitting easier

## Options to read_csv
Check out "?read_csv" in console for more info
If there are blank rows at the beginning of the file, can skip them
Can also make R import values and desginate them as NA 
```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
litters_df = read_csv("./data/FAS_litters.csv", na = c("", "NA", ".", 999))
```


## Other file formats

Read in an EXCEL file
Look at "?read_excel" in console for more info
```{r}
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
```


Read in a SAS file 

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
```

## Comparison with Base R

What about "read.csv" --> This is built into R but Jeff would recomment we use "read_csv"

```{r}
litters_base = read.csv("./data/FAS_litters.csv")
litters_readr = read_csv("./data/FAS_litters.csv")
```
Printed differences!
USE "read_csv"!!!

## Exporting data

Export the mlb sub-table. 

```{r}
write_csv(mlb_df, "./data/mlb_subtable.csv")
```






